---
title: "Homeplay TERGM"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remedy::remedy_opts$set(name = paste0("btergm_homeplay", "_"))
```


# Introduction

For this homeplay we are going to use the `friendship` data from the `SNA4DSData` 
package. You probably need to update the `SNA4DSData` 
package for it, through


`remotes::install_github("SNAnalyst/SNA4DSData")`


The load the data and the `btergm` package as follows:

```{r , eval = FALSE}
data(friendship, package = "SNA4DSData")
library(btergm)
```

This loads the dataset into your workspace. The data consist of the 
friendship network of 26 pupils in a Dutch school class measured at four time 
points along with several demographic and behavioral covariates. 


Check the number of vertices per network as follows

```{r, eval = FALSE}
sapply(friendship, snafun::count_vertices)
```


# Model I

Fit the first model on this dataset. 
Although the vertex covariates speak for themselves, this is what they represent:

- idegsqrt = the square root of the vertex indegree at that point in time

- odegsqrt = the square root of the vertex outdegree at that point in time

- sex: gender, 1 = girl, 2 = boy

We start with the model below (`model.01`). Run it now.

```{r , eval = FALSE}
model.01 <- btergm::btergm(friendship ~ edges + mutual + ttriple +
                     transitiveties + ctriple + nodeicov("idegsqrt") +
                     nodeicov("odegsqrt") + nodeocov("odegsqrt") +
                     nodeofactor("sex") + nodeifactor("sex") + nodematch("sex") +
                     edgecov(primary), R = 100)

snafun::stat_plot_gof_as_btergm(model.01, 
                                btergm_statistics = c(esp, dsp, geodesic, deg, triad.undirected, rocpr))
```

The last line shows how you can easily specify the statistics you want to check for 
the goodness of fit. If you leave out the `btergm_statistics` argument, you'll 
only get the default statistics (ie. shared partners, geodesic distances, degree, roc/pr).

After you run the model:
* step 1: interpret the findings of the model--which effect are (not) statistically significant and does that make sense?
(hint: use `btergm::summary(model.01)`.
* Step 2: What do you think of the model's goodness of fit?

# Model II

Model I didn't include any history. Lets add some now. 

* Extend model I with "delayed reciprocity." Use the default arguments.

* Also, add a stability memory effect, again using the defaults.

Call this new model `model.02`. 

Hint, use 

```{r , eval = FALSE}
help("tergm-terms")
```

to find out how to specify the various time effects. 
Of course, you can also use the lecture slides to get you started.


```{r , echo = FALSE, eval = FALSE}
model.02 <- btergm::btergm(friendship ~ edges + mutual + ttriple +
                     transitiveties + ctriple + nodeicov("idegsqrt") + nodeicov("odegsqrt") +
                     nodeocov("odegsqrt") + nodeofactor("sex") + nodeifactor("sex") +
                     nodematch("sex") + edgecov(primary) + delrecip +
                     memory(type = "stability"), R = 100)
```


After you run the model:
* step 1: interpret the findings of the model--which effect are (not) statistically significant and does that make sense?
* Step 2: What do you think of the model's goodness of fit?


# Model III

The final model is actually the same model that you just formulated. 
But this time, we are going to only run it on the first three networks and use 
the fitted model to predict the fourth. 
This is a useful way to check the fit of our model through out-of-sample prediction.

STEP 1

Fit Model II, but now use `friendship[1:3]` as the set of dependent networks, 
rather than `friendship` (which would fit the model on all networks). 
For the rest, don't change anything in the model.
Name the model `model.03`.


```{r , echo = FALSE, eval = FALSE}
model.03 <- btergm::btergm(friendship[1:3] ~ edges + mutual + ttriple +
                     transitiveties + ctriple + nodeicov("idegsqrt") + nodeicov("odegsqrt") +
                     nodeocov("odegsqrt") + nodeofactor("sex") + nodeifactor("sex") +
                     nodematch("sex") + edgecov(primary) + delrecip() +
                     memory(type = "stability"), R = 100)
```

OK, so how do we predict and know how well the prediction does? 
We can employ the `gof` function to simulate 100 networks from the model and compare
them to the actually observed network at $t = 4$. 

For the prediction, we use the argument `target = friendship[[4]]` to tell the 
`gof` function to compare the simulated networks to the omitted
and observed fourth time step. To simulate new networks, one needs to supply the covariates
for the coeﬃcients, including the temporal statistics. We do this by specifying the formula
again and adding index [3:4] to the left-hand side of the formula. This computes the
relevant covariates for the fourth time step, partly by making use of information contained in
the third time step (in the case of the temporal variables). Using the `coef` argument, we tell
the `gof` function to use the estimated coeﬃcients from the model.

As an example, suppose that you model.03 looked as follows:

```{r btergm_homeplay_001, eval = FALSE}
model.03 <- btergm::btergm(friendship[1:3] ~ edges + A + B + C, R = 100)
```

Then, you predict the next network by:

```{r, echo = TRUE , eval = FALSE}
gof.03 <- btergm::gof(model.03, nsim = 100, target = friendship[[4]],
              formula = friendship[3:4] ~ A + B + C, coef = coef(model.03),
              statistics = c(esp, dsp, geodesic, deg, triad.undirected, rocpr))
```

Do this now for the actual model you fitted.
(if you have a slow computer, you can reduce `nsim`, but this should run within a 
few minutes on most laptops).


```{r, echo = TRUE , eval = FALSE}
gof.03 <- btergm::gof(model.03, nsim = 100, target = friendship[[4]],
              formula = friendship[3:4] ~ edges + mutual + ttriple +
                transitiveties + ctriple + nodeicov("idegsqrt") +
                nodeicov("odegsqrt") + nodeocov("odegsqrt") + nodeofactor("sex") +
                nodeifactor("sex") + nodematch("sex") + edgecov(primary) +
                delrecip + memory(type = "stability"), coef = coef(model.03),
              statistics = c(esp, dsp, geodesic, deg, triad.undirected, rocpr))
```

To extract the area uder the ROC and PR curves, use:

```{r , eval = FALSE}
gof.03$`Tie prediction`$auc.roc
gof.03$`Tie prediction`$auc.pr
```


For your information, you can also simulate a set of networks without automatically 
extracting the `gof` like you did above.

The following command simulates ten new networks based on the coeﬃcients stored 
in Model.03 and for time step 4. 
The resulting ten networks are stored in a list. 
This would be the code to do that (but you don't need to do it for this homeplay, 
this is just so you know how to do it, in case you would like to):

```{r , eval = FALSE}
nw <- simulate(model.03, nsim = 10, index = 4)
```

