---
title: "Homeplay for (intro to) Statistical Network Models"
output: 
  html_document: 
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
      number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  include = FALSE,
  eval = FALSE
)

data(fifa2006, package = "SNA4DSData")
data(fifa2015, package = "SNA4DSData")
data(louis, package = "SNA4DSData")
```

## Louisiana

Open the dataset `louis` from the `SNA4DSData` package. 

The network describes the spatial adjacency (ie: do two counties share a border)
among the counties. The data were collected just before the 1960 elections in Louisiana.

You will see that the vertices have several attributes:

* democratic : the proportion of support in the county for democratic presidential candidate Kennedy.
* Intercept : a vector of ones, needed to get an intercept in the NAM model
* perc_black : percentage black in the county
* perc_cath: percentage catholic
* perc_urban: percentage urban
* black_political_equality: measure of black political equality in the county

Our interest in this exercise is in modeling the extent to which political
preference seem to cluster between adjacent counties. This makes the
variable 'democratic' the dependent variable for this analysis.
We also want to see if counties that have mutually equivalent locations in
the spatial network tend to vote similar.

1. Create a Weight matrix based on the adjacency matrix. 
Make sure to row normalize the matrix.<br>
(NOTE: you can use the cheatsheet to find the 
appropriate functions, in this case the function to create 
the adjacency matrix).

1. Run an autocorrelation model where democratic preference 
is regressed on "perc_black", "perc_cath", "perc_urban", and "black_political_equality."
<br>(do not forget the intercept)

1. Interpret the findings

```{r hp06_remedy01}
data(louis, package = "SNA4DSData")
attrs <- snafun::extract_all_vertex_attributes(louis)
attrs <- attrs[, - which(colnames(attrs) %in% c("Intercept", "vertex.names"))]

adj <- snafun::to_matrix(louis)
w_adj <- adj/rowSums(adj)
mod <- snafun::stat_nam(democratic ~ ., data = attrs, W = w_adj)
summary(mod)
```


NOTE: Remember that you get access to an attribute using the 
`snafun::extract_vertex_attribute`
function. So `snafun::extract_vertex_attribute(louis, "perc_black")` gives you the
vector with the variable `perc_black`.
You can also extract all of the vertex attribute values into a data.frame using the
function `snafun::extract_all_vertex_attributes`, so here you would use
`snafun::extract_all_vertex_attributes(louis)` to get a data.frame with all of
the vertex attributes values of the louis network.

It is advisable to remove the attributes "vertex.names" and "Intercept" from the 
resulting attributes data.frame, so they don't end up in the model.


## FIFA

(NOTE: this assignment might take some time to run--I suggest you make some tea
or coffee to enjoy while you wait for the analysis to finish, if you have a
slow-ish laptop. If you have a slow laptop, limit yourself to 100 or 200 replications.)

For the second analysis, load the datasets `fifa2006` and `fifa2015`
from `SNA4DSData`.

The vertices in the fifa2006 network are the members of the standing committees
in FIFA. The edges represent joint membership. So, an edge between i and j
means that both i and j were member of the same standing committee inside
FIFA in 2006.

The fifa2015 represents the committee membership network for the year 2015.

Let's look at the division of power inside Fifa in these two years.

First, transitivity. At its core, transitivity reflects the tendency that
"a friend of my friend is my friend". That means that if you and I are on a
committee together, and I am also on a committee with Claudia, then you would
be tempted to start working on a committee with Claudia as well.

1. Calculate the transitivity for both of the networks.

```{r hp06_remedy03}
snafun::g_transitivity(fifa2006)
snafun::g_transitivity(fifa2015)
```

2. Run a test to check if the transitivities of the two fifa networks are statistically
significantly high or low. Control for network size and density.


```{r hp06_remedy04}
# condition on the density
trans_f <- function(x, directed = FALSE) {
  x <- snafun::fix_cug_input(x, directed = directed)
  snafun::g_transitivity(x)
}


system.time(
cug_trans_06 <- sna::cug.test(fifa2006, mode = "graph", 
                              FUN = trans_f, 
                              cmode = "edges", reps = 1000)
)
cug_trans_06


system.time(
cug_trans_15 <- sna::cug.test(fifa2015, mode = "graph", 
                              FUN = trans_f, 
                              cmode = "edges", reps = 1000)
)
cug_trans_15
```

3. You could argue that the previous test is insufficient, since committee 
membership
is very skewed among the fifa members most people have few memberships and some
have very many.<br>
Try: `hist(snafun::v_degree(fifa2006))` or `hist(snafun::v_degree(fifa2015))`.<br><br>
Therefore, now run a test to check if the transitivities of the two fifa
networks are statistically
significantly high or low, controling for network size **and** degree distribution.
<br>


```{r fifahist}
hist(snafun::v_degree(fifa2006),
     breaks = 20,
     main = "Degree distribution of Fifa in 2006", 
     xlab = "Number of shared committees")
hist(snafun::v_degree(fifa2015), 
     breaks = 20,
     main = "Degree distribution of Fifa in 2015", 
     xlab = "Number of shared committees")
```


```{r hp06_remedy05}
# condition on the dyad census
betw_f <- function(x, directed = FALSE) {  # note: directed = FALSE!
  x <- snafun::fix_cug_input(x, directed = directed)
  snafun::g_centralize(x, measure = "betweenness", directed = directed)$centralization
}

system.time(
cug_betw_06_ds <- sna::cug.test(fifa2006, 
                                mode = "graph", FUN = betw_f, 
                                cmode = "dyad.census", reps = 200)
)
cug_betw_06_ds

system.time(
cug_betw_15_ds <- sna::cug.test(fifa2015, 
                                mode = "graph", FUN = betw_f, 
                                cmode = "dyad.census", reps = 200)
)
cug_betw_015_ds

```

4. Interpret your findings.

5. Let's also check if the distribution of advantageous positions in the
administrative network is exceptionally high. In this case, calculate some
centralization score (e.g., betweenness centralization) and compare that score
with networks of similar size+density or of similar degree distribution.
You can do this for both networks, or just take one, depending on how much
time you are willing to wait for a result.<br><br>


```{r}
system.time(cug_betw_06 <- sna::cug.test(fifa2006,
                                         FUN = sna::centralization,
                                         FUN.arg=list(FUN = sna::betweenness),
                                         mode="graph",
                                         reps = 200,
                                         cmode="edges"))

system.time(cug_betw_15 <- sna::cug.test(fifa2015,
                                          sna::centralization,
                                          FUN.arg=list(FUN = sna::betweenness),
                                          mode="graph",
                                          reps = 200,
                                          cmode="edges"))

system.time(cug_betw_06_ds <- sna::cug.test(fifa2006,
                                            FUN = sna::centralization,
                                            FUN.arg=list(FUN = sna::betweenness),
                                            mode="graph",
                                            reps = 200,
                                            cmode="dyad.census"))

system.time(cug_betw_15_ds <- sna::cug.test(fifa2015,
                                             sna::centralization,
                                             FUN.arg=list(FUN = sna::betweenness),
                                             mode="graph",
                                             reps = 200,
                                             cmode="dyad.census"))
```



6. interpret your findings.




