---
title: "Homeplay 02 -- Using R for SNA"
author: "Roger Leenders"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this little homeplay, you are going to familiarize yourself with the SNA4DS 
cheatsheet and will experience what makes the `snafun` so bloody awesome.
You'll also gain some experience with using multiple packages as part of 
your analysis pipeline.

You can find the cheatsheet here: [SNAFUN site](https://snanalyst.github.io/SNAfun/).

You will be asked to calculate several measures. 
We know that you may not know what these measures mean yet, but that is OK. 
The exercise is meant to help you find the specific functions quickly and apply 
them correctly. We'll dive into their actual meaning in the very next lecture 
of this course.

1. Load the dataset `elegans` from the `SNA4DSData` package

```{r include=FALSE}
data(elegans, package = "SNA4DSData")
elegans
snafun::print(elegans)
snafun::is_igraph(elegans)

snafun::is_directed(elegans)
snafun::list_edge_attributes(elegans)
snafun::list_vertex_attributes(elegans)

snafun::count_vertices(elegans)
snafun::count_edges(elegans)

snafun::extract_vertex_names(elegans)

#dyad census
snafun::count_dyads(elegans)

snafun::g_reciprocity(elegans)
snafun::g_transitivity(elegans)
snafun::g_density(elegans)
```

For the next steps (2 - 9), use the `snafun` package only. Even if you can see the 
answer already from the prinout of the object itself, still use `snafun` functions 
to verify the result (this will help you become more fluent at the programming part).

2. Print the data object;

1. Extract the number of vertices and the number of edges in the object;

1. Check that the object is of class `igraph` ;

1. Check that the object is directed;

1. List the vertex attributes and the edge attributes;

1. Extract the vertex names;

1. Calculate the dyad census;

1. Calculate the reciprocity, transitivity, and density of the `elegans` object.

1. Now do all of the steps above (from printing the object through calculating its 
density) using the `igraph` package only.

```{r, include = FALSE}
elegans
igraph::is_igraph(elegans)
igraph::is_directed(elegans)
igraph::edge_attr_names(elegans)
igraph::vertex_attr_names(elegans)
igraph::vcount(elegans)
igraph::ecount(elegans)
igraph::V(elegans)$name
igraph::dyad_census(elegans)
igraph::reciprocity(elegans)
igraph::transitivity(elegans, type = "global")
igraph::edge_density(elegans)
```

1. Turn `elegans` into an object of class  `network`, using the appropriate function 
from the`snafun` package. Call the network object `elegans_n`.

1. Do all of the steps from printing the object through calculating its 
density using the `network` and `sna` packages (use `elegans_n` 
and check for a `network` object in step 4).

```{r include=FALSE}
# network
elegans_n <- snafun::to_network(elegans)
network::is.network(elegans_n)
elegans_n
network::is.directed(elegans_n)
network::list.edge.attributes(elegans_n)
network::list.vertex.attributes(elegans_n)
network::network.edgecount(elegans_n)
network::network.size(elegans_n)
network::get.vertex.attribute(elegans_n, "vertex.names")
sna::dyad.census(elegans_n)
sna::grecip(elegans_n, measure = "edgewise")
sna::gtrans(elegans_n, mode = "digraph", measure = "weak")
sna::gden(elegans_n)
```


Did you manage to do that? I knew you could. Please notice the (reasonable)
consistency of the `snafun` package API. 
The `igraph` package is fairly consistent too, but the 
functions you needed for `elegans_n` are much less consistent in their input, 
output, and function naming. The `network` package is all over the place.
